---
title: "R Notebook"
output: html_notebook
---
# Purpose:
Create a method of evaluating potential properties for the profitability of different investment strategies based on the length of time the user is looking to invest for.

# Goals:
1. Create an estimation method for the profitability of a property for a given investment strategy
2. Create estimates for multiple investment strategies and rank them form most to worst
3. Use this method of analysis to identify exceptionally profitable properties that are currently on the market.


load up the packages necessary for some good data exploration. Download the data set from St. Louis county housing.
```{r include = FALSE}
library(ggplot2)
library(tidyr)
library(dplyr)
library(data.table)
library(haven)
library(jsonlite)
setwd('/home/tracy/R/data_incubator_challenge/')
df <- fread("stlouis.csv")

```

```{r}
unique(df$PHYSCITY)
```

```{r} 
# filter so only private owners in Duluth are left for initial methods
df <- df %>%
  filter(Ownership == "Private")%>%
  filter(PHYSCITY == "Duluth" | PHYSCITY== "DULUTH MN") %>%
  select(OBJECTID, ACREAGE, TPCLS1, LAND_EST, BUILDING, EstTotalValue, TAXABLE_BUILDING, TAXABLE_LAND_VALUE, TaxableMarketValue, NET_TAX)
```

# Long-term rental calculation
Determine monthly cash flow.
Expenses:
-Tax = given in the dataset
-Insurance= 120 needs a model. Right now I'm just locking in at $120 arbitrarily
-Utilities= 200 needs a lookup. Right now locking in at $200
-HOA = 0 in Duluth
-Lawn snow = 40 make it user defined eventually if they want to do it themselves or hire out
-Repairs=100 estimate
-Capital expenditures= 100 savings for long term projects (black swans)
-Manager= 60 User defined if they want someone to collect for them and deal with minor issues
-PMI = 0 for down payments over 20%
Income:
-Rent= user defined (create a function that evaluates rent potential)
-Others= TBD (maybe laundry etc.)

```{r}
Insurance <- 120
Utilities <- 200
HOA <- 0
Lawn_snow <- 40
Repairs <- 100
Capex <- 100
Manager <- 60
PMI <- 0
Rent <- 15000

df <- df %>%
  mutate(Cash_flow= Rent - (NET_TAX/12 + Insurance + Utilities + HOA + Lawn_snow+ Repairs + Capex + Manager + PMI)) %>%
  mutate(Yearly_flow = Cash_flow * 12)
```

# Determine ROI
Factors:
-Down payment = user defined
-Closing = 2-5% property value
-Initial renovations = 0 user defined
```{r}
Down = 0.2 * df$EstTotalValue
Closing = 0.05 * df$EstTotalValue
Renovs  = 0

df <- df %>%
  mutate(First_year_ROI = Yearly_flow/(Down + Closing + Renovs))
```

